---
title: "cmc regression"
author: "Danielle Zhang"
date: "22/01/2018"
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    code_folding: show
---

## Import data

```{r, message=FALSE}
library(tidyverse)

x = read_csv("data/AtoPandTcmc.csv") %>% 
  janitor::clean_names() %>% mutate(date = lubridate::as_date(timestamp))

x2= x %>% filter(retail == 'pre') %>% 
  dplyr::select(cmc, hunter_l, hunter_a, hunter_b, nix_a, nix_b, nix_l, retail)

```


## Modelling with Lab

### Nix L vs Hunter L


```{r}
# l model using hunter l to predict nix l

x2 %>% ggplot(aes(x = hunter_l, y= nix_l)) + 
  geom_point() + 
  geom_smooth(method = "lm") + 
  geom_abline() + 
  coord_cartesian(xlim = c(10,60),ylim = c(10,60))

l_model = lm(nix_l ~ hunter_l, data = x2)
summary(l_model)

# l model using nix l to predict hunter l

x2 %>% ggplot(aes(y = hunter_l, x= nix_l)) + 
  geom_point() + 
  geom_smooth(method = "lm") + 
  geom_abline() + 
  coord_cartesian(xlim = c(10,60),ylim = c(10,60))


l_model = lm(hunter_l ~ nix_l, data = x2)
summary(l_model)
```


### Simple linear regressions of cmc against L a b

cmc with hunter_l-b
```{r}
# consumer meat colour score

cmc_model = lm(cmc~ hunter_l+ hunter_a + hunter_b, data = x2)
summary(cmc_model)

p1 = x2 %>% ggplot(aes(y = cmc, x= hunter_l)) + 
  geom_point() + 
  geom_smooth(method = "lm") 

p2 = x2 %>% ggplot(aes(y = cmc, x= hunter_a)) + 
  geom_point() + 
  geom_smooth(method = "lm") 

p3 = x2 %>% ggplot(aes(y = cmc, x= hunter_b)) + 
  geom_point() + 
  geom_smooth(method = "lm") 

p1
p2
p3

```

```{r}
#par(mfrow = c(2,2))
library(ggfortify)
autoplot(cmc_model)
#plot(cmc_model)
```

cmc hunter model with interactive predictors

```{r}
cmc_model2 = lm(cmc~ hunter_l*hunter_a*hunter_b, data = x2)
summary(cmc_model2)

cmc_model3 = lm(cmc~ hunter_l + hunter_a*hunter_b, data = x2)
summary(cmc_model3)

anova(cmc_model, cmc_model2)

# par(mfrow = c(2,2))
##autoplot(cmc_model3)
# plot(cmc_model2)
# plot(cmc_model3)
```


cmc model with nix lab 

```{r}
cmc_model4 = lm(cmc~ nix_l + nix_a + nix_b, data = x2)
summary(cmc_model4)

cmc_model5 = lm(cmc~ nix_l * nix_a * nix_b, data = x2)
summary(cmc_model5)

cmc_model5b = lm(cmc~ nix_l * nix_a + nix_l * nix_b + nix_a * nix_b, data = x2)
summary(cmc_model5b)

anova(cmc_model4, cmc_model5)
anova(cmc_model5b, cmc_model5)

```


non-linear cmc model with hunter lab

```{r}
cmc_model7 = lm(cmc~ hunter_l + hunter_a+ hunter_b + I(hunter_l ^2), data = x2)
summary(cmc_model7)
plot(cmc_model7)
```

```{r}
cmc_model8 = lm(cmc~ hunter_l*hunter_a*hunter_b +  I(hunter_l ^2) + I(hunter_a ^2) + I(hunter_b ^2) , data = x2)
summary(cmc_model8)
plot(cmc_model8)

```


```{r}
cmc_model9a = x2 %>% 
  dplyr::select(cmc, hunter_l, hunter_a, hunter_b) %>% 
  na.omit() %>% 
  lm(cmc~ hunter_b*hunter_a*hunter_l  + poly(hunter_l,4) + poly(hunter_a,4) + poly(hunter_b,4) , data = .)

cmc_model9b = x2 %>% 
  dplyr::select(cmc, hunter_l, hunter_a, hunter_b) %>% 
  na.omit() %>% 
  lm(cmc~hunter_b*hunter_a*hunter_l+ poly(hunter_l,3) + poly(hunter_a,3) + poly(hunter_b,3) , data = .)

summary(cmc_model9b)

anova(cmc_model9b, cmc_model9a)

cmc_model9c = x2 %>% 
  dplyr::select(cmc, hunter_l, hunter_a, hunter_b) %>% 
  na.omit() %>% 
  lm(cmc~ hunter_b*hunter_a*hunter_l + poly(hunter_l,2) + poly(hunter_a,2) + poly(hunter_b,2), data = .)

summary(cmc_model9c)

anova(cmc_model9c, cmc_model9b)


```

filtered data well picked up, go back and filter to pre so we've only got one unique cmc value for each sample.

```{r}
x3 = x2 %>% filter(retail == 'pre')
x3 %>% ggplot(aes(y = cmc, x= hunter_l)) + 
  geom_point() + 
  geom_smooth(method = "lm")

x3 %>% ggplot(aes(y = cmc, x= hunter_a)) + 
  geom_point() + 
  geom_smooth(method = "lm")

x3 %>% ggplot(aes(y = cmc, x= hunter_b)) + 
  geom_point() + 
  geom_smooth(method = "lm")

```

## CMYK regressions

pre, nix cmyk regression

```{r}
x4 = x %>% filter(retail == 'pre') %>% dplyr::select(eqsref, cmc, nix_cyan, nix_magenta, nix_yellow, nix_key)

x4 %>% ggplot(aes(y = cmc, x= nix_cyan)) + 
  geom_point() + 
  geom_smooth(method = "lm") 

x4 %>% ggplot(aes(y = cmc, x= nix_magenta)) + 
  geom_point() + 
  geom_smooth(method = "lm") 

x4 %>% ggplot(aes(y = cmc, x= nix_yellow)) + 
  geom_point() + 
  geom_smooth(method = "lm") 

x4 %>% ggplot(aes(y = cmc, x= nix_key)) + 
  geom_point() + 
  geom_smooth(method = "lm") 
```


```{r}
cmyk_model = lm(cmc~ nix_cyan + nix_magenta + nix_yellow + nix_key, data = x4)
summary(cmyk_model)

```

```{r}
cmyk_model2 = lm(cmc~ nix_cyan * nix_magenta + nix_cyan * nix_yellow + nix_cyan * nix_key + nix_magenta * nix_yellow + nix_magenta * nix_key + nix_yellow * nix_key, data = x4)
summary(cmyk_model2)

anova(cmyk_model, cmyk_model2)
```

```{r}
cmyk_model3 = lm(cmc~ nix_cyan * nix_magenta * nix_yellow * nix_key - nix_cyan : nix_magenta : nix_yellow : nix_key, data = x4)
summary(cmyk_model3)
```

```{r}
cmyk_model4 = lm(cmc~ nix_cyan * nix_magenta * nix_yellow * nix_key, data = x4)
summary(cmyk_model4)

```

```{r}
anova(cmyk_model, cmyk_model4)
```

```{r}
cmyk_model5 = lm(cmc~ poly(nix_cyan,2) + poly(nix_magenta,2) + poly(nix_yellow,2) + poly(nix_key,2), data = x4 %>% 
  dplyr::select(cmc, nix_cyan, nix_magenta, nix_yellow, nix_key) %>% na.omit())
summary(cmyk_model5)
```

```{r}
cmyk_model6 = lm(cmc ~ I(nix_cyan ^2) + I(nix_magenta ^2) + I(nix_yellow ^2) + I(nix_key ^2) + nix_cyan * nix_magenta * nix_yellow * nix_key, data = x4)
summary(cmyk_model6)
```

## RGB regressions

pre, nix rgb regression

```{r}
x5 = x %>% filter(retail == 'pre') %>% dplyr::select(cmc, nix_red, nix_green, nix_blue)

x5 %>% ggplot(aes(y = cmc, x= nix_red)) + 
  geom_point() + 
  geom_smooth(method = "lm") 

x5 %>% ggplot(aes(y = cmc, x= nix_green)) + 
  geom_point() + 
  geom_smooth(method = "lm") 

x5 %>% ggplot(aes(y = cmc, x= nix_blue)) + 
  geom_point() + 
  geom_smooth(method = "lm") 
```

```{r}
rgb_model = lm(cmc ~ nix_red + nix_green + nix_blue, data = x5)
summary(rgb_model)
```

```{r}
rgb_model2 = lm(cmc ~ nix_red * nix_green * nix_blue - nix_red : nix_green : nix_blue, data = x5)
summary(rgb_model2)

anova(rgb_model, rgb_model2)
```

```{r}
rgb_model3 = lm(cmc ~ nix_red * nix_green * nix_blue, data = x5)
summary(rgb_model3)

anova(rgb_model, rgb_model3)
```

```{r}
rgb_model4 = lm(cmc ~ nix_red + nix_green + nix_blue + I(nix_red^2) + I(nix_green ^2) + I(nix_blue^2), data = x5)
summary(rgb_model4)
anova(rgb_model, rgb_model4)
```

```{r}
rgb_model5 = lm(cmc ~ nix_red * nix_green * nix_blue + I(nix_red^2) + I(nix_green ^2) + I(nix_blue^2), data = x5)
summary(rgb_model5)
anova(rgb_model, rgb_model5)
```

```{r}
rgb_model6 = lm(cmc ~ nix_red + nix_green + nix_blue + nix_red : nix_green + nix_red : nix_blue, data = x5)
summary(rgb_model6)
anova(rgb_model, rgb_model6)

```

```{r}
rgb_model7 = lm(cmc~ I(nix_red ^-1) + I(nix_green ^-1) + I(nix_blue ^-1), data = x5)
summary(rgb_model7)
anova(rgb_model, rgb_model7)
```

```{r}
rgb_model8 = lm(cmc~ I(nix_red ^-1) + I(nix_green ^-1) + I(nix_blue ^-1) + nix_red + nix_green + nix_blue, data = x5)
summary(rgb_model8)
anova(rgb_model, rgb_model8)
```


```{r}
lab_model = lm(cmc ~ hunter_l + hunter_a + hunter_b + I(hunter_l ^-1) + I(hunter_a ^-1) + I(hunter_b ^-1), data = x2)
summary(lab_model)
anova(cmc_model, lab_model)
```

```{r}
lab_model2 = lm(cmc ~ I(hunter_l ^-1) + I(hunter_a ^-1) + I(hunter_b ^-1) + hunter_l * hunter_a * hunter_b - hunter_l : hunter_a : hunter_b, data = x2)
summary(lab_model2)
anova(cmc_model, lab_model2)
```

```{r}
lab_model3 = lm(cmc ~ I(hunter_l ^-1) + I(hunter_a ^-1) + I(hunter_b ^-1) + hunter_l * hunter_a * hunter_b , data = x2)
summary(lab_model3)
anova(lab_model2, lab_model3)

anova(cmc_model2, lab_model2)

```

```{r}
lab_model4 = lm(cmc ~ I(hunter_l ^-1) + I(hunter_a ^-1) + I(hunter_b ^-1) + hunter_l * hunter_a * hunter_b - hunter_l : hunter_a : hunter_b + I(hunter_l ^2) + I(hunter_a ^2) + I(hunter_b ^2), data = x2)
summary(lab_model4)
anova(lab_model2, lab_model4)
```


